---
title: "Global Fertility rates and GDP relationship"
format: html
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

## Average fertility rates by continent  

Fertility rates has been much better in the western nations due to high GDP and healthcare.

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("maps")
library(maps)

#creating objects based on the csv files
unicef_metadata <- read_csv("/cloud/project/unicef_metadata_posit.csv")
unicef_indicator_2 <- read_csv("/cloud/project/unicef_indicator_2_posit.csv")
data_right_3 <- read_csv("/cloud/project/data_right_3.csv")
indicator_data3 <- merge(unicef_indicator_2,data_right_3, by = "country")
overall_sheet <- merge(indicator_data3,unicef_metadata,  by.x = c("country","time_period"), by.y = c("country","year"))

#creating average fertility value of each continent
average_by_continent <- indicator_data3 %>%
  group_by(continent) %>%
  summarise(avg_obs_value = mean(obs_value, na.rm = TRUE))


#plotting average fertility of each continent in a bar graph
average_by_continent %>%
  ggplot(aes(x = continent, y = avg_obs_value)) +
  geom_bar(stat = "identity", fill = "#187832") +
  labs(title = "Fertility rate Average by continent",
       x = "Continent",
       y = "Fertility Rate") +
  guides(fill = FALSE) + 
  geom_text(aes(label = round(avg_obs_value,2)), vjust = -0.5, color = "black", size = 3.5) +
  theme_minimal()

```


## Africa's fertility rates over the years

The minimum and maximum fertility rates of each African country across the timeline.

```{r}
# time series for fertility rate

continent_africa_fertility <- overall_sheet %>%
  filter(continent == "Africa")
  
  ggplot(data = continent_africa_fertility, aes(x = time_period, y = obs_value, group = country, color = country)) +
    geom_line() +
    xlab("Year") +
    ylab("Observation Value") +
    theme_minimal() + 
    guides(color = FALSE)
    


```


## Relationship between GDP and Fertility rate

Focusing further on Africa, the most populous country, Nigeria's fertilty rate has improved with rise in GDP.

```{r}
#scatter plot for fertility rate
  nigeria_data <- continent_africa_fertility %>%
    filter(country == "Nigeria") 
  
  nigeria_data %>%
    ggplot(aes(x = `GDP per capita (constant 2015 US$)`, y = obs_value, color = obs_value) )+
    geom_point(alpha=2.0) +
    scale_color_gradient(low = "#98e2eb", high = "#ed8582", name = "obs_value") +
    labs(x = "GDP per capta",
         y = "fertility Rate") +
    theme_minimal() + 
    theme(legend.position = "hide")
```


## Fertility rate: a global view

In the global population growth reached a peak in 1963 with an annual growth rate of 2.2%; but since then, world population growth has halved. So, this year is a crucial year to review. 

```{r}
 # mapping the world map of fertility rate
  world_map <- map_data("world")
  
 fertility_data_1963 <- overall_sheet  %>%
    filter(time_period == 1963) %>%
    merge(world_map, by.x = c("country"), by.y = c("region"), all.x = TRUE)
  
  
 
  ggplot() +
    geom_polygon(data = fertility_data_1963, aes(x = long, y = lat, group = group, fill = obs_value)) +
    scale_fill_gradient(low = "#e80c0c", high = "#37c43e", name = "obs_value", limits = c(0, 10), breaks = seq(0, 10, by = 1)) +
    theme_minimal() +
    theme(legend.position = "hide")
  
```